Segmentation of Electron Microscopy Images for Connectomics

### 2018-05-01 papers/quals2018/0_introduction.tex
quals: added method and evaluation from eccv submission

The field of connectomics is concerned with the reconstruction of the wiring diagram of the mammalian brain.
In order to achieve this ambitious goal, we need images at nanometer resolution to see the structure of the neurons and the synaptic connections between them.
Recent advances in electron microscopy acquisition techniques enable a throughput of a terabyte of image data every hour~\cite{richard2016imaging}.
Manual reconstruction of this image data is simply infeasible and requires automatic segmentation methods.
These automatic methods label every voxel of the image with a 64-bit integer value where two voxels have the same label only if they belong to the same neuron.
As the field progresses, our goal is to reconstruct a cubic millimeter of brain which will result in over 18 petabytes of segmentation data alone~\cite{suissa2016automatic}.

Three major challenges arise as the image datasets scale to a petabyte and more. 
First, the segmentation methods must be accurate enough to extract the proper connections between neurons.
In particular, the dendritic spines \TODO{check,CITE} which contain the postsynaptic density can be only \TODO{XX} nanometers thick.
Second, automatic segmentation must be fast -- ideally matching the throughput of the electron microscope~\cite{haehn2017scalable}.
The current state-of-the-art automatic reconstruction pipelines are either too slow for such large datasets or too inaccurate at scale.
Third, the label volumes require $8\times$ the number of bytes as the already massive image volumes.
Storing such data is expensive and transferring the data is slow. 
To cut costs and delays, we need compression methods to reduce data sizes.

>>>>

One of the critical components of connectomics--the field concerned with reconstructing the wiring diagram of the brain at nanometer resolution--is automatic segmentation of electron microscopy (EM) images of the brain.
With recent advances in EM acquisition techniques, neuroscientists can now generate a terabyte of EM image data every hour~\cite{richard2016imaging}.
These images are typically anisotropic with $4$nm resolution in the $xy$ plane with $30-40$nm between slices. 
At this resolution, we can see all of the axon terminals and dendritic spines, and the synaptic connections between them.
Accurate reconstruction of every neuron paired with synapse detection will enable us create a true graphical representation of the brain and further our understanding of the underlying circuitry.
These segmentation, or label, volumes assign an integer label to every voxel where voxels have the same label only if they belong to the same neuron.

The first complete reconstruction of animal brain occurred in the 1980s with an ambitious manual effort of the C. elegans worm\TODO{CITE}. 
This species has \TODO{300} neurons and manual labeling required \TODO{13} years of intensive work.
More recent research focuses on Drosophila flies~\cite{jovanic2016competitive}\TODO{CITE JANELIA}, rodents~\cite{richard2016imaging,suissa2016automatic}, and even humans\TODO{CITE}. 
With an EM throughput of one terabyte per hour, neuroscientists can image a cubic millimeter of data (one petabyte) in six months. 
At this scale, the manual reconstruction efforts of the 80s are simply infeasible.
Thus, researchers have turned to the deep learning revolution to provide automatic solutions to the segmentation problem.

Automatic reconstruction techniques need to be fast, scalable, and accurate.
Ideally, image acquisition is the bottleneck of the connectomics pipeline which requires reconstruction efforts of a terabyte per hour~\cite{haehn2017scalable}.
We can achieve this throughput with parallelization among several GPUs with fast reconstruction techniques~\cite{funke2017deep,parag2017anisotropic}.
One of the issues with these automatic techniques is that they use only local context in decision making and are agnostic about the underlying biological system.
Thus, these methods make errors at scale and currently require human proofreading and other bulky error correction techniques~\cite{haehn2017guided,error_correction_using_CNN}.

We propose a novel region merging framework that takes as input and oversegmentation of an EM volume.
Our method imposes both local and global biological constrains onto the output segmentation to more close match the underlying structure of neuronal processes.
Additionally, our method is independent of image resolution and acquisition parameters, enabling its application to isotropic and anisotropic data without retraining.
Images generated by electron microscopes often differ in appearance because of variations in staining techniques~\cite{briggman2012volume}.
By removing the dependence of our algorithm on the input images, we greatly reduce the need for additional costly ground truth data for each new stack of images.

### 2018-05-01 papers/quals2018/1_related_work.tex
quals: added method and evaluation from eccv submission

\paragraph{Automatic Segmentation.}
A significant amount of connectomics research considers the problem of extracting segmentation information at the voxel level of EM images.
First, intermediate representations like boundary, affinity or binary segmentation maps are generated from the voxels.
Random forests with hand-designed features~\cite{kaynig2015large}, or 2D and 3D convolutional networks produce boundary probabilities~\cite{seymour2016rhoananet,ronneberger2015u,bogovic2013learned,ciresan2012deep,jain2010boundary,amelio_segmentation}.
Often, the affinity between each voxel and its six neighbors are used~\cite{lee2015recursive,parag2017anisotropic,lee2017superhuman,cciccek20163d,turaga2010convolutional}. 
The 3D U-Net architecture has become popular~\cite{cciccek20163d} and the MALIS cost function is specifically designed to re-weight affinity predictions by their contribution to the segmentation error~\cite{briggman2009maximin}.
More recently, flood-filling networks~\cite{januszewski2016flood} produce binary segmentations from raw pixels with a recurrent convolutional network at a high computational cost.
Orthogonal to the representation, model averaging~\cite{zeng2017deepem3d} and data augmentation~\cite{lee2017superhuman} methods can further improve the performance, where Lee et al.~\cite{lee2017superhuman} surpass the estimated human accuracy on the SNEMI3D dataset.

@@

Region merging methods can be categorized by the similarity metric between adjacent segments and the merging strategy.
For the similarity metric, Lee et al. and Funke et al. rely solely on the accuracy of the predicted affinities and define the metric as the mean affinity between segments~\cite{lee2017superhuman,funke2017deep}.
Classification-based methods generate the probability to merge two segments from handcrafted~\cite{seymour2016rhoananet,nunez2014graph,parag2017anisotropic,zlateski2015image,10.1371/journal.pone.0125825,jain2011learning} and learned features~\cite{bogovic2013learned}. 
For the merging strategy, most methods use variants of hierarchical agglomeration ~\cite{seymour2016rhoananet,nunez2014graph,parag2017anisotropic,zlateski2015image,10.1371/journal.pone.0125825} to greedily merge a pair of regions at a time.
Jain et al. formulates the agglomeration problem as a reinforcement learning problem ~\cite{jain2011learning} and Pape et al. present a scalable multicut algorithm to partition superpixels with global optimization~\cite{beier2017multicut}.

Additional research builds on top of these region-based methods to correct errors in the segmentation. This can be done either using human proofreading~\cite{haehn2014design,haehn2017guided,mojo2} or automatically~\cite{rolnick2017morphological,error_correction_using_CNN}. In both cases, available methods are pixel-based and do not include global biological constraints into the decision making process. Our method can be used as input to any existing error correction framework.

\paragraph{Compression.}

The literature currently lacks efficient compression of label volumes. 
General-purpose compression schemes~\cite{deutsch1996zlib,ziv1978compression,lehmann2016liblzf,oberhumer2005lzo,welch1984technique,seward1998bzip2,pavlov2007lzma,vandevenne2016zopfli,google2016brotli,collet2016smaller} are not optimized for this data.
In this paper, we exploit the typical characteristics of label volumes such as large invariant regions without natural relationship between label values. 
These properties render 2D image compression schemes inadequate since they rely on frequency reduction (using e.g., wavelet or discrete cosine transform) and value prediction of pixels based on local context (differential pulse-code modulation)~\cite{roelofs1999png,skodras2001jpeg}. 
Color space optimization strategies in video codecs~\cite{aimar2005x264} also have no effect on label volumes, even though the spatial properties of a segmentation stack (\textit{z}-axis) are similar to the temporal properties of video data (time-axis). 
A compression scheme designed specifically for label volumes is part of the visualization software Neuroglancer~\cite{google2016compressed}. 
This method exploits segmentation homogeneity by creating small blocks with $N$ labels and reducing local entropy to $\log_2{N}$ per pixel. 
Lookup tables then decode the values $[0,N)$ to the original 64-bit labels. We compare the Neuroglancer scheme with our method.

>>>>

\begin{figure*}[t!]
	\centering
	\includegraphics[width=0.75\linewidth]{./figures/teaser_v4.png}
	\caption{Our framework uses both geometric and topological biological constraints for region merging. We (a) generate a simplified graph from the skeleton representation of the input segmentation, (b) train a classifier to learn the edge weight based on the shape of the segment connection, and (c) perform graph partitioning with acyclic constraints.}
	\label{fig:teaser_pipeline}
\end{figure*}

A significant amount of connectomics research considers the problem of extracting segmentation information at the voxel level of EM images.
First, intermediate representations like boundary, affinity or binary segmentation maps are generated from the voxels.
Random forests with hand-designed features~\cite{kaynig2015large}, or 2D and 3D convolutional networks produce boundary probabilities~\cite{seymour2016rhoananet,ronneberger2015u,bogovic2013learned,ciresan2012deep,jain2010boundary,amelio_segmentation}.
Often, the affinity between each voxel and its six neighbors are used~\cite{lee2015recursive,parag2017anisotropic,lee2017superhuman,cciccek20163d,turaga2010convolutional}. 
The 3D U-Net architecture has become popular~\cite{cciccek20163d} and the MALIS cost function is specifically designed to re-weight affinity predictions by their contribution to the segmentation error~\cite{briggman2009maximin}.
More recently, flood-filling networks~\cite{januszewski2016flood} produce binary segmentations from raw pixels with a recurrent convolutional network at a high computational cost.
Orthogonal to the representation, model averaging~\cite{zeng2017deepem3d} and data augmentation~\cite{lee2017superhuman} methods can further improve the performance, where Lee et al.~\cite{lee2017superhuman} surpass the estimated human accuracy on the SNEMI3D dataset.

@@

Region merging methods can be categorized by the similarity metric between adjacent segments and the merging strategy.
For the similarity metric, Lee et al. and Funke et al. rely solely on the accuracy of the predicted affinities and define the metric as the mean affinity between segments~\cite{funke2017deep,lee2017superhuman}.
Classification-based methods generate the probability to merge two segments from handcrafted~\cite{jain2011learning,seymour2016rhoananet,nunez2014graph,10.1371/journal.pone.0125825,parag2017anisotropic,zlateski2015image} and learned features~\cite{bogovic2013learned}. 
For the merging strategy, most methods use variants of hierarchical agglomeration ~\cite{seymour2016rhoananet,nunez2014graph,10.1371/journal.pone.0125825,parag2017anisotropic,zlateski2015image} to greedily merge a pair of regions at a time.
Jain et al. formulates the agglomeration problem as a reinforcement learning problem ~\cite{jain2011learning} and Pape et al. present a scalable multicut algorithm to partition superpixels with global optimization~\cite{beier2017multicut}.

Additional research builds on top of these region-based methods to correct errors in the segmentation. This can be done either using human proofreading~\cite{haehn2017guided,haehn2014design,mojo2} or automatically~\cite{rolnick2017morphological,error_correction_using_CNN}. In both cases, available methods are pixel-based and do not include global biological constraints into the decision making process. Our method can be used as input to any existing error correction framework.
